---
title: "TV news mentions of White House list of attacks"
author: "Andrew Ba Tran"
date: "2/11/2017"
output: html_document
---

White House officials released a list of 78 attacks they claimed the media did not cover following President Donald Trump's accusations at MacDill Air Force Base in Florida on Monday.

"It's gotten to a point where it's not even reported, and in many cases the very, very dishonest press doesn't even want to report it," said President Trump to the troops gathered there.

An analysis of news channel transcripts collected by the Internet Archive Television News Archive shows that the international attacks were indeed mentioned, with the exception of a couple non-fatal incidents in Copenhagen, Denmark.


```{r setup,  warning=F, message=F}
library(newsflash)
library(ggalt) # github version
library(hrbrmisc) # github only
library(tidyverse)
library(ggplot2)
library(scales)
library(lubridate)
library(stringr)


wh_list <- read.csv("72attacks.csv", stringsAsFactors=F)
wh_list$keyword <- wh_list$Location
wh_list$keyword <- gsub(",.*", "", wh_list$keyword)
wh_list$keyword <- gsub("\\(.*", "", wh_list$keyword)
wh_list$keyword <- str_to_title(wh_list$keyword)
wh_list$keyword2 <- gsub(".*,", "", wh_list$Location)
wh_list$keyword2 <- str_to_title(wh_list$keyword2)
wh_list$month <- gsub(", .*", "", wh_list$Date)
wh_list$year <- gsub(".*,", "", wh_list$Date)
wh_list$start <- paste0(wh_list$month, " 1, ", wh_list$year)
wh_list$start <- mdy(wh_list$start)
wh_list$end <- wh_list$start+30

wh_list$ID<-seq.int(nrow(wh_list))

stations <- list_networks(widget=FALSE)
colnames(stations) <- c("keyword", "station", "dates")
stations$dates <- NULL
```

## Each incident

```{r each, fig.height=5, fig.width=8, warning=F, message=F}

for (i in 1:nrow(wh_list)) {
starts <- wh_list$start[i]
ends <- wh_list$end[i]

queries <- map2(starts, ends, function(x, y) {
  query_tv(wh_list$keyword[i], "attack", timespan="custom", start_date=x, end_date=y)
})

queries2 <- map2(starts, ends, function(x, y) {
  query_tv(wh_list$keyword2[i], "attack", timespan="custom", start_date=x, end_date=y)
})

if (!is.null(queries[[1]])) {
queries_timeline <- map_df(queries, "timeline")


queries_df <- queries_timeline %>%
  group_by(station) %>%
  summarize(count=sum(value))

stat_m<- left_join(stations, queries_df)
stat_m <- subset(stat_m, !is.na(count))
stat_m$ID <- i
stat_m$total <- sum(stat_m$count)

if (i == 1) {
  stations_m <- stat_m
} else {
  stations_m <- rbind(stations_m, stat_m)
}

count(queries_timeline, station, wt=value, sort=TRUE) %>%
  mutate(pct=n/sum(n), pct_lab=sprintf("%s (%s)", scales::comma(n), scales::percent(pct)),
         station=factor(station, levels=rev(station))) -> timeline_df

gg <- ggplot(timeline_df, aes(y=station, x=pct))
gg <- gg + geom_lollipop(point.colour="steelblue", point.size=3, horizontal=TRUE)
gg <- gg + geom_text(label=timeline_df$pct_lab, hjust=-.3)
gg <- gg + scale_x_continuous(labels=percent,
                              breaks=seq(0, 1, by=0.2), limits=c(0, 1))
gg <- gg + labs(x=NULL, y=NULL,
                title=paste0("Percent of all sentences on TV channels mentioning ", wh_list$keyword[i]),
                subtitle=paste0("Search parameters: Primary [", wh_list$keyword[i], "]; Context: [attack]; n=", sum(stat_m$total)),
                caption="Source: Internet Archive Television News Archive via #rstats newsflash package")
gg <- gg + theme_minimal(base_family="Arial Narrow")
gg <- gg + theme(panel.grid.major.y=element_blank())
gg <- gg + theme(panel.grid.minor=element_blank())
gg <- gg + theme(axis.line.y=element_line(color="#2b2b2b", size=0.15))
gg <- gg + theme(axis.text.y=element_text(margin=margin(r=-5, l=0)))
gg <- gg + theme(plot.margin=unit(rep(30, 4), "pt"))
gg <- gg + theme(plot.title=element_text(face="bold"))
gg <- gg + theme(plot.subtitle=element_text(margin=margin(b=10)))
gg <- gg + theme(plot.caption=element_text(size=8, margin=margin(t=10)))
print(gg)

} else if ((is.null(queries[[1]]) & !is.null(queries2[[1]]))) {
  queries_timeline <- map_df(queries2, "timeline")
  
  
  queries_df <- queries_timeline %>%
    group_by(station) %>%
    summarize(count=sum(value))
  
  stat_m<- left_join(stations, queries_df)
  stat_m <- subset(stat_m, !is.na(count))
  stat_m$ID <- i
  stat_m$total <- sum(stat_m$count)
  
  if (i == 1) {
    stations_m <- stat_m
  } else {
    stations_m <- rbind(stations_m, stat_m)
  }
  
  count(queries_timeline, station, wt=value, sort=TRUE) %>%
    mutate(pct=n/sum(n), pct_lab=sprintf("%s (%s)", scales::comma(n), scales::percent(pct)),
           station=factor(station, levels=rev(station))) -> timeline_df
  
  gg <- ggplot(timeline_df, aes(y=station, x=pct))
  gg <- gg + geom_lollipop(point.colour="steelblue", point.size=3, horizontal=TRUE)
  gg <- gg + geom_text(label=timeline_df$pct_lab, hjust=-.3)
  gg <- gg + scale_x_continuous(labels=percent,
                                breaks=seq(0, 1, by=0.2), limits=c(0, 1))
  gg <- gg + labs(x=NULL, y=NULL,
                  title=paste0("Percent of all sentences on TV channels mentioning ", wh_list$keyword[i]),
                  subtitle=paste0("Search parameters: Primary [", wh_list$keyword[i], "]; Context: [attack]; n=", sum(stat_m$total)),
                  caption="Source: Internet Archive Television News Archive via #rstats newsflash package")
  gg <- gg + theme_minimal(base_family="Arial Narrow")
  gg <- gg + theme(panel.grid.major.y=element_blank())
  gg <- gg + theme(panel.grid.minor=element_blank())
  gg <- gg + theme(axis.line.y=element_line(color="#2b2b2b", size=0.15))
  gg <- gg + theme(axis.text.y=element_text(margin=margin(r=-5, l=0)))
  gg <- gg + theme(plot.margin=unit(rep(30, 4), "pt"))
  gg <- gg + theme(plot.title=element_text(face="bold"))
  gg <- gg + theme(plot.subtitle=element_text(margin=margin(b=10)))
  gg <- gg + theme(plot.caption=element_text(size=8, margin=margin(t=10)))
  print(gg)
  
} else {
stat_temp <- data.frame(keyword="", station="", count=0, ID=i, total=0)
stations_m <- rbind(stations_m, stat_temp)
}
}

```

## Total incidents

```{r each_all, fig.height=12, fig.width=10, warning=F, message=F}

wh_list_narrow <- select(wh_list, Location, Date, ID)
wh_list_narrow <- left_join(stations_m, wh_list_narrow, by="ID")
wh_list_narrow$loc_id <- paste0(wh_list_narrow$Location, " (", wh_list_narrow$Date, ")")
wh_list_short <- wh_list_narrow %>%
  group_by(loc_id) %>%
  filter(row_number()==1)
wh_list_short$keyword <- NULL
wh_list_short$station <- NULL
wh_list_short$count <- NULL


write.csv(wh_list_narrow, "wh_list_narrow.csv")

gg <- ggplot(data=wh_list_narrow, aes(x=reorder(loc_id, -ID), y=count))
gg <- gg + geom_bar(stat="identity", aes(fill=station))
gg <- gg + geom_text(data=wh_list_short, aes(y=total, label=total), size=3, hjust=-.3)
gg <- gg + coord_flip()
gg <- gg + scale_y_continuous(labels=comma, limits=c(0, 7000))
gg <- gg + labs(x=NULL, y=NULL,
                title="Mentions of international attacks on TV news",
                subtitle="Search parameters: Primary [Location]; Context: [attack]",
                caption="Source: Internet Archive Television News Archive via #rstats newsflash package")
gg <- gg + theme_minimal(base_family="Arial Narrow")
gg <- gg + theme(panel.grid.major.y=element_blank())
gg <- gg + theme(panel.grid.minor=element_blank())
gg <- gg + theme(axis.line.y=element_line(color="#2b2b2b", size=0.15))
gg <- gg + theme(axis.text.y=element_text(margin=margin(r=-5, l=0)))
gg <- gg + theme(plot.margin=unit(rep(30, 4), "pt"))
gg <- gg + theme(plot.title=element_text(face="bold"))
gg <- gg + theme(plot.subtitle=element_text(margin=margin(b=10)))
gg <- gg + theme(plot.caption=element_text(size=8, margin=margin(t=10)))
gg

```

```{r fox, fig.height=12, fig.width=8, warning=F, message=F}
wh_list_narrow2 <- as.data.frame(wh_list_narrow)
gg <- ggplot(data=subset(wh_list_narrow, station=="FOX News"), aes(x=reorder(loc_id, -ID), y=count))
gg <- gg + geom_bar(stat="identity", aes(fill=station))
gg <- gg + geom_text(data=subset(wh_list_narrow, station=="FOX News"), aes(y=count, label=count), size=3, hjust=-.3)
gg <- gg + coord_flip()
gg <- gg + scale_y_continuous(labels=comma, limits=c(0, 7000))
gg <- gg + labs(x=NULL, y=NULL,
                title="Mentions of international attacks on FOX News",
                subtitle="Search parameters: Primary [Location]; Context: [attack]",
                caption="Source: Internet Archive Television News Archive via #rstats newsflash package")
gg <- gg + theme_minimal(base_family="Arial Narrow")
gg <- gg + theme(panel.grid.major.y=element_blank())
gg <- gg + theme(panel.grid.minor=element_blank())
gg <- gg + theme(axis.line.y=element_line(color="#2b2b2b", size=0.15))
gg <- gg + theme(axis.text.y=element_text(margin=margin(r=-5, l=0)))
gg <- gg + theme(plot.margin=unit(rep(30, 4), "pt"))
gg <- gg + theme(plot.title=element_text(face="bold"))
gg <- gg + theme(plot.subtitle=element_text(margin=margin(b=10)))
gg <- gg + theme(plot.caption=element_text(size=8, margin=margin(t=10)))
gg
```